Hierarchical Dispatching and Overriding

### 2017-09-18 NewVersion/sections/Formalization.tex
.
reply marco
refine
remove StaticInvocation
More proof.

\begin{figure*}[t]
\saveSpaceFig
\begin{displaymath}
\begin{array}{l}
\begin{array}{llrl}
\text{Interfaces}   & IL & \Coloneqq & \interface{I}{I}{M} \\
\text{Methods}      & M  & \Coloneqq & \method{I}{m}{I_x}{x}{J}{e}  \mid
									   \absmethod{I}{m}{I_x}{x}{J} \\
\text{Expressions}  & e  & \Coloneqq & x \mid
e.m(\overline{e}) \mid
\new{I} \mid
e.\staAnno{I}{J}m(\overline{e}) \mid
e.I?m(e) \; \mid \; \angl{I}e \\
\text{Context}      & \Gamma & \Coloneqq & \overline{x}:\overline{I} \\
\text{Values}       & v & \Coloneqq & \angl{I} \new{J} \\
%%\\
%%\text{Interface names} & I, J, K & & \\
%%\text{Method names} & m & & \\
%%\text{Variable names} & x & &
\end{array}
\end{array}
\end{displaymath}
\caption{Syntax of \name{}.}\label{fig:syntax}
\saveSpaceFig
\end{figure*}

@@

\begin{figure*}[t]
\saveSpaceFig
\begin{mathpar}
	\framebox{$ I <: J $} \hspace{.5in} \subid \\
	\subtrans \hspace{.5in} \subextends \\
	\framebox{$ \judgeewf \Gamma {e:I} $} \hspace{.5in}
	\tvar \\
	\tinvk \\
	\tpathinvk \\
	% \tsuperinvk \\
	\tstaticinvk \\
	\tnew \\
	\tmethod \\
	\tabsmethod \\
	\tintf
\end{mathpar}
\caption{Typing and subtyping rules.}\label{fig:typingrules}
\saveSpaceFig
\end{figure*}

@@

\begin{figure*}[t]
\saveSpaceFig
\begin{mathpar}
	\sinvk \\
	\spathinvk \\
	% \ssuperinvk \\
	\sstaticinvk \\
	\creceiver \hspace{.5in}
	\cpathreceiver \\
	\cargs \\
	\cpathargs \\
	% \csuperargs \\
	\cstaticargs \\
	\cstatictype \\
	\cfreduce \\
	\cannoreduce
\end{mathpar}
\caption{Small-step semantics.}\label{fig:smallstep}
\saveSpaceFig
\end{figure*}

@@

\begin{theorem}[Progress]~\label{theorem_progress}
$\textit{Suppose } e \textit{ is a well-typed expression } $ \\
\begin{enumerate}
\item If $e$ includes $\left(\angl{J}\emph{\kwnew}\;I()\right).m(\overline{\angl{I_e}} \; \overline{e})$ as a subexpression,
    then $ \emph{\mbody}(m, I, J) = (J', \overline{I_x} \; \overline{x}, I'_E \; e_0) $ and
         $\num{\overline{x}} = \num{\overline{e}}$ for some $\overline{I_x}, \overline{x}, I_e'$ and $e_0$.
\item If $e$ includes $\left(\angl{J}\emph{\kwnew}\;I()\right).K::m(\overline{\angl{I_e}} \overline{e})$ as a subexpression,
    then $ \emph{\mbody}(m, I, K) = (J', \overline{I_x} \; \overline{x}, I'_E \; e_0) $ and 
         $\num{\overline{x}} = \num{\overline{e}}$ for some $\overline{I_x}, \overline{x}, I_e'$ and $e_0$.
\item If $e$ includes $\emph{\kwsuper}.K::m(\overline{\angl{I_e}} \overline{e})$ as a subexpression,
    then $ \emph{\mbody}(m, K, K) = (J', \overline{I_x} \; \overline{x}, I'_E \; e_0) $ and 
         $\num{\overline{x}} = \num{\overline{e}}$ for some $\overline{I_x}, \overline{x}, I_e'$ and $e_0$.
\end{enumerate}
\end{theorem}
\begin{proof}
See Appendix~\ref{appendix_proof}.
\end{proof}

>>>>

\begin{figure*}[t]
\saveSpaceFig
\begin{displaymath}
\begin{array}{l}
\begin{array}{llrl}
\text{Interfaces}   & IL & \Coloneqq & \interface{I}{I}{M} \\
\text{Methods}      & M  & \Coloneqq & \method{I}{m}{I_x}{x}{J}{e}  \mid
									   \absmethod{I}{m}{I_x}{x}{J} \\
\text{Expressions}  & e  & \Coloneqq & x \mid
e.m(\overline{e}) \mid
\new{I} \mid
e.I?m(e) \; \mid \; \angl{I}e \\
\text{Context}      & \Gamma & \Coloneqq & \overline{x}:\overline{I} \\
\text{Values}       & v & \Coloneqq & \angl{I} \new{J} \\
%%\\
%%\text{Interface names} & I, J, K & & \\
%%\text{Method names} & m & & \\
%%\text{Variable names} & x & &
\end{array}
\end{array}
\end{displaymath}
\caption{Syntax of \name{}.}\label{fig:syntax}
\saveSpaceFig
\end{figure*}

@@

\begin{figure*}[t]
\saveSpaceFig
\begin{mathpar}
	\framebox{$ I <: J $} \hspace{.5in} \subid \\
	\subtrans \hspace{.5in} \subextends \\
	\framebox{$ \judgeewf \Gamma {e:I} $} \hspace{.5in}
	\tvar \\
	\tinvk \\
	\tpathinvk \\
	% \tsuperinvk \\
	% \tstaticinvk  \\
	\tnew \\
	\tmethod \\
	\tabsmethod \\
	\tintf \\
	\tanno
\end{mathpar}
\label{fig:typingrules}
\saveSpaceFig
\end{figure*}

@@

\begin{figure*}[t]
\saveSpaceFig
\begin{mathpar}
	\sinvk \\
	\spathinvk \\
	% \ssuperinvk \\
	% \sstaticinvk \\
	\creceiver \hspace{.5in}
	\cpathreceiver \\
	\cargs \\
	\cpathargs \\
	% \csuperargs \\
	% \cstaticargs \\
	\cstatictype \\
	\cfreduce \\
	\cannoreduce
\end{mathpar}
\caption{Small-step semantics.}\label{fig:smallstep}
\saveSpaceFig
\end{figure*}

@@

\begin{theorem}[Progress]~\label{theorem_progress}
$\textit{Suppose } e \textit{ is a well-typed expression, if } e \textit{ includes} $

\noindent $\left(\angl{J}\emph{\kwnew}\;I()\right).K?m(\overline{v}) \textit{ as a subexpression, then } \emph{\mbody}(m, I, K) = (I_0, \overline{I_x} \; \overline{x}, I_e\; e_0) $ $\textit{and } \num{\overline{x}} = \num{\overline{e}} \textit{ for some } \overline{I_x}, \overline{x}, I_e \textit{ and } e_0.$
\end{theorem}
\begin{proof}
See Appendix~\ref{appendix_proof}.
\end{proof}

### 2017-09-18 NewVersion/sections/TypeSafety.tex
More proof.

\begin{lemma}[Method Type Preservation]~\label{lemma2}
$\textit{If } \mtype(m, I) = \overline{I_x} \rightarrow I_e, \textit{ then } \mtype(m, J) = \overline{I_x} \rightarrow I_e 
\textit{ for all } J \subtype I.
$
\begin{proof}~\\
Straight induction on the derivation of $J \subtype I$. Whether $m$ is defined in $J$ or not, $\mtype(m, J)$ should 
be the same as $\mtype(m, K)$ where $J \; \kwextends \; K \{...\}$.
\end{proof}
\end{lemma}

\textcolor{red}{
\begin{lemma}[Method Type Preservation]~\label{lemma2_new}
$\textit{If } \mbody(m, I_0, J_0) = (K, \overline{I_x}\ \overline{x}, I_e\ e)$, $\textit{ and }
\mbody(m, I'_0, J_0) = (K', \overline{I'_x}\ \overline{x'}, I'_e\ e') \textit{ for some }I'_0 \subtype I_0, \textit{ then }
\overline{I'_x} = \overline{I_x}$ $\textit{and } I'_e = I_e.$
\end{lemma}}

@@

\noindent \textbf{Case Invk.}
$ e = e_0.m(\overline{e}) \quad
  \judgeewf {\Gamma, \overline{x}:\overline{I_x}} {e_0 : I_0} $
$$ mtype(m, I_0) = \overline{I_e} \rightarrow J $$
$$ \judgeewf {\Gamma, \overline{x}:\overline{I_x}} {\overline{e}:\overline{I}} \quad
    \overline{I} \subtype \overline{I_e} $$
By induction hypothesis, there are some $I_0'$ and $\overline{I_e'}$ such that
    $$ \judgeewf {\Gamma} {[\overline{y}/\overline{x}]e_0 : I_0'} \quad 
        I_0' \subtype I_0 $$
    $$ \judgeewf {\Gamma} {[\overline{y}/\overline{x}]\overline{e} : \overline{I_e'}} \quad  
        \overline{I_e'} \subtype \overline{I}$$    
By lemma~\ref{lemma2}, 
    $mtype(m, I_0') = \overline{I_e} \rightarrow J$.

@@

\noindent \textbf{Case SuperInvk.}
$ e = \kwsuper.I::m(\overline{e}) $ \\
Suppose $\judgeewf {\Gamma} {\kwthis : I_0}$, the following proof should be similar as case Var.

@@

\subsubsection{Proof for Theorem~\ref{theorem_subject}}
\begin{proof} ~\\
\noindent\textcolor{red}{\textbf{Case PathInvk.} Let} \haoyuan{where to use lemma 1}
  $$\textcolor{red}{e = (\angl{J}\new{I}).K?m(\overline{v}) \quad \judgeewf {\Gamma} {e : I_e}}$$
  $$\textcolor{red}{e' = \angl{I_e}[\overline{\angl{I_x}} \overline{v}/\overline{x}, \angl{I_0}\new{I}/\kwthis]e_0 \quad \judgeewf {\Gamma} {e' : I'_e}}$$
  $$\textcolor{red}{\mbody(m, I, K) = (I_0, \overline{I_x} \; \overline{x}, I_e \; e_0)}$$
\textcolor{red}{Since $e$ is well typed, $\angl{J}\new{I}$ is well typed, and}
  $$\textcolor{red}{\judgeewf {\Gamma} {\angl{J}\new{I} : I}}$$
\textcolor{red}{By \textsc{(T-Anno)}, we have $I \subtype J$. Also, by \textsc{(T-PathInvk)}, $\judgeewf {\Gamma} {\overline{v} : \overline{I_v} \subtype \overline{I_x}}$.}

\noindent \textcolor{red}{So the first substitution preserves typing, On the other hand, .... the second one also preserves typing. Hence $I'_e \subtype$ type of $e_0 \subtype I_e$.}

\noindent \textbf{Case Invk.} 
let \[ e = \angl{J}\new I.m(\overline{\angl{I_e}}\overline{e}) \] 
Suppose \[ \mbody(m, I, J) = (J', \overline{I_x} \; \overline{x}, I_{e_0} \; e_0) \] 
then \[ e' =  [\overline{\angl{I_x}} \overline{e}/\overline{x}, \; \angl{J}\new I/\kwthis ] e_0 \] 
By rules \textsc{(T-New)} and \textsc{(T-Invk)}, 
  \[ \judgeewf \Gamma {\new I:I} \quad 
     \mtype(m, I, J) = \overline{I_x} \rightarrow I_{e_0} \quad 
     \judgeewf \Gamma {\overline{e} : \overline{I_e'}} \quad
     \overline{I_e'} \subtype \overline{I_x} \quad
     \textit{, for some } \; \overline{I_e'}
  \]
By Lemma~\ref{lemma0},
    \[
    \judgeewf {\Gamma, \overline{x}:\overline{I_x}, \kwthis:J_0} {e_0:I_f} \textit{, for some } J \subtype J_0 \textit{ and } I_f \subtype I_{e_0}
    \]
By Lemma~\ref{lemma1},
    \[
    \judgeewf {\Gamma} {[\overline{\angl{I_x}} \overline{e}/\overline{x}, \; \angl{J}\new I/\kwthis ] e_0  :  I_g} \textit{, for some } I_g \subtype I_f 
    \]
So $I_g <: I_{e_0}$, finally just let $I' = I_g$.

@@

\subsubsection{Proof for Theorem~\ref{theorem_progress}}
\begin{proof}~\\
\noindent \textbf{Case 1.}
Given that $e$ is well-typed, by rule \textsc{(T-Invk)}, $\mtype(m, J)$ is defined.
By rule \textsc{(T-Intf)} for interface $I$,
    \[ \forall J >: I, m, \mtype(m, J) \text{ is defined } \Rightarrow \mbody(m, I, J) \text{ is defined } \]
So straightforwardly, we get $ \mbody(m, I, J) = (J', \overline{I_x} \; \overline{x}, I'_E \; e_0) $ and
         $\num{\overline{x}} = \num{\overline{e}}$ for some $J', \overline{I_x}, \overline{x}, I_e'$ and $e_0$

\noindent \textbf{Case 2 and 3.} Proof for case 2 and 3 is similar to case 1, and we omit it here.
\end{proof}

>>>>

\textcolor{red}{
\begin{lemma}[Method Type Preservation]~\label{lemma2_new}
$\textit{If } \mbody(m, I_0, J_0) = (K, \overline{I_x}\ \_, I_e\ \_)$, $\mbody(m, I'_0, J_0) = (K', \overline{I'_x}\ \_, I'_e\ \_) \textit{ for some }I'_0 \subtype I_0, \textit{ then }
\overline{I'_x} = \overline{I_x}$ $\textit{and } I'_e = I_e.$
\end{lemma}}
\begin{proof}

Trivial for the case $I'_0 = I_0$. And since the lemma has transitivity on subtyping, it suffices to prove the following cases and rely on induction, where we can suppose that $I'_0$ directly extends $I_0$.

Suppose that $$\mostSpecific(m, I_0, J_0) = \{I\}$$ $$\mostSpecificOverride(m, I_0, I) = \{K\}$$

\textbf{(1) The body of $I'_0$ does not contain $m$.} Trivial because $\mbody(m, I_0, J_0) = \mbody(m, I'_0, J_0)$. \haoyuan{Non-trivial. m may be inherited from other super types. TODO.}

\textbf{(2) $m \in \overrideSet(I'_0, I'_0)$.} In that case $I'_0$ defines an original method $m$.
 So
$$\mostSpecific(m, I'_0, J_0) = \{I'_0\}$$ $$\mostSpecificOverride(m, I'_0, I'_0) = \{I'_0\}$$
Hence $\mbody(m, I'_0, J_0)$ returns the original $m$ from $I'_0$. By $\textsc{(T-Intf)}$, we know that $\canOverride(m, I'_0, I)$ holds, which means that $I'_0$ has the same type as $I$ for $m$. On the other hand, from $\textsc{(T-Method)}$ and $\textsc{(T-AbsMethod)}$, the type of $m$ is also preserved between $I$ and $K$ (method override needs to keep the type). Finally $\mbody(m, I'_0, J_0)$ has the same type as $\mbody(m, I_0, J_0)$.

\textbf{(3) $m \in \overrideSet(I'_0, I)$.} Similar to (2), $I$ and $K$ have the same type for $m$. Moreover, $\textsc{(T-Method)}$ and $\textsc{(T-AbsMethod)}$ also ensure that $I'_0$ and $I$ have the same type for $m$. \haoyuan{But findOrigin may no longer be I.}

\textbf{(4) $m \in \overrideSet(I'_0, J)$ for some other $J$.} Same as (1). Such an $m$ does not affect the result of $\mbody$.
%Straight induction on the derivation of $J \subtype I$. Whether $m$ is defined in $J$ or not, $\mtype(m, J)$ should 
%be the same as $\mtype(m, K)$ where $J \; \kwextends \; K \{...\}$.
\end{proof}

@@

\begin{comment}
\noindent \textbf{Case Invk.}
$ e = e_0.m(\overline{e}) \quad
  \judgeewf {\Gamma, \overline{x}:\overline{I_x}} {e_0 : I_0} $
$$ mtype(m, I_0) = \overline{I_e} \rightarrow J $$
$$ \judgeewf {\Gamma, \overline{x}:\overline{I_x}} {\overline{e}:\overline{I}} \quad
    \overline{I} \subtype \overline{I_e} $$
By induction hypothesis, there are some $I_0'$ and $\overline{I_e'}$ such that
    $$ \judgeewf {\Gamma} {[\overline{y}/\overline{x}]e_0 : I_0'} \quad 
        I_0' \subtype I_0 $$
    $$ \judgeewf {\Gamma} {[\overline{y}/\overline{x}]\overline{e} : \overline{I_e'}} \quad  
        \overline{I_e'} \subtype \overline{I}$$    
By lemma~\ref{lemma2}, 
    $mtype(m, I_0') = \overline{I_e} \rightarrow J$.

@@

\noindent \textbf{Case SuperInvk.}
$ e = \kwsuper.I::m(\overline{e}) $ \\
Suppose $\judgeewf {\Gamma} {\kwthis : I_0}$, the following proof should be similar as case Var.
\end{comment}

@@

\subsubsection{Proof for Theorem~\ref{theorem_subject}}
\begin{proof} ~\\
\noindent\textcolor{red}{\textbf{Case Invk.} TODO. By t-invk and t-intf.}

\noindent\textcolor{red}{\textbf{Case PathInvk.} Let}
  $$\textcolor{red}{e = (\angl{J}\new{I}).K?m(\overline{v}) \quad \judgeewf {\Gamma} {e : I_e}}$$
  $$\textcolor{red}{e' = \angl{I_e}[\overline{\angl{I_x}} \overline{v}/\overline{x}, \angl{I_0}\new{I}/\kwthis]e_0}$$
  $$\textcolor{red}{\mbody(m, I, K) = (I_0, \overline{I_x} \; \overline{x}, I_e \; e_0)}$$
\textcolor{red}{Since $\mbody(m, I, K)$ is defined, the definition of $\mbody$ ensures that $I \subtype I_0$. And since $e$ is well typed, by \textsc{(T-PathInvk)},}
  $$\textcolor{red}{\judgeewf {\Gamma} {\overline{v} : \overline{I_v} \subtype \overline{I_x}}}$$

\noindent \textcolor{red}{By \textsc{(T-New)}, \textsc{(T-Anno)} and lemma~\ref{lemma0},}
  $$\textcolor{red}{\judgeewf {\Gamma} {\angl{I_0}\new{I} : I} \quad \judgeewf {\Gamma, \overline{x} : \overline{I_x}, \kwthis : I_0} {e_0 : I_{e_0} \subtype I_e}}$$

\noindent \textcolor{red}{Hence by lemma~\ref{lemma1}, the substitution preserves typing, thus}
  $$\textcolor{red}{\judgeewf {\Gamma} {[\overline{\angl{I_x}} \overline{v}/\overline{x}, \angl{I_0}\new{I}/\kwthis]e_0 : I'_{e_0} \subtype I_{e_0} \subtype I_e}}$$
\noindent \textcolor{red}{By \textsc{(T-Anno)}, $\judgeewf {\Gamma} {e' : I'_{e_0} \subtype I_e}$.}

\begin{comment}
\noindent \textbf{Case Invk.} 
let \[ e = \angl{J}\new I.m(\overline{\angl{I_e}}\overline{e}) \] 
Suppose \[ \mbody(m, I, J) = (J', \overline{I_x} \; \overline{x}, I_{e_0} \; e_0) \] 
then \[ e' =  [\overline{\angl{I_x}} \overline{e}/\overline{x}, \; \angl{J}\new I/\kwthis ] e_0 \] 
By rules \textsc{(T-New)} and \textsc{(T-Invk)}, 
  \[ \judgeewf \Gamma {\new I:I} \quad 
     \mtype(m, I, J) = \overline{I_x} \rightarrow I_{e_0} \quad 
     \judgeewf \Gamma {\overline{e} : \overline{I_e'}} \quad
     \overline{I_e'} \subtype \overline{I_x} \quad
     \textit{, for some } \; \overline{I_e'}
  \]
By Lemma~\ref{lemma0},
    \[
    \judgeewf {\Gamma, \overline{x}:\overline{I_x}, \kwthis:J_0} {e_0:I_f} \textit{, for some } J \subtype J_0 \textit{ and } I_f \subtype I_{e_0}
    \]
By Lemma~\ref{lemma1},
    \[
    \judgeewf {\Gamma} {[\overline{\angl{I_x}} \overline{e}/\overline{x}, \; \angl{J}\new I/\kwthis ] e_0  :  I_g} \textit{, for some } I_g \subtype I_f 
    \]
So $I_g <: I_{e_0}$, finally just let $I' = I_g$.

@@

\end{comment}
\end{proof}

\subsubsection{Proof for Theorem~\ref{theorem_progress}}
\begin{proof}~\\
\noindent \textcolor{red}{TODO. t-pathinvk. t-new. remember canInstantiate() to ensure $e_0$ is not empty.}
\end{proof}
